### 2025-03-09 TSUSC/conclusion.tex

\section{Conclusion}
			\label{sec:conclusion}
This paper proposed offline and online algorithms for the EV charging scheduling problem under fractional and integral revenue models in an adaptive charging network (ACN). The problem is different, and more challenging than the existing single station EV charging scheduling problems since it requires respecting the aggregate peak charging demand of the ACN. 
%More importantly, the problem is online in nature in the sense that the EVs arrive in time and the scheduler has no information regarding future arrivals. 
As the notable contributions, our proposed online algorithm for fractional revenue model achieves constant competitive ratio of $2$. Moreover, the offline integral algorithm achieves a theoretical bound on the optimality gap and approximates the optimum by $92\%$, on average in experiments. As a future work, we plan to study the problem under posted pricing mechanism where the charging station publishes the unit price of the power (which can be varied over time) and the users can accept or reject the offer.
\revv{Another interesting future direction is to tackle the EV charging scheduling in market-based scenarios where the prices change over time.}
%As a future work, we plan to analyze the truthfulness of the algorithms to make sure that it is a dominant strategy for the EV owners to declare their true valuations and availabilities.
%\vspace{-4mm}

%\input{intro.tex}
%\input{rel.tex}
%\input{sys.tex}		
%\input{fraction.sol.tex}		
%\input{integral.sol.tex}
%\input{exp.tex}
%%%\vspace{0.1mm}
%\input{conc.tex}
%\newpage
%ss
%\newpage
%\pagestyle{empty}